---
title: "DiseaseDiva"
author: "Us"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background 
A team of 73 researchers from all over the United States investigated clinical genomics of advanced prostate cancer. The aim of this research was to facilitate precision medicine create a prospective 



Cleaning Our Data

```{r}
#Load data downloaded from 
#rawdat <- read.csv("/cloud/project/data/diseasedivadata.csv")
rawdat<- read.csv("~/Documents/Dartmouth/Fall 2021/QBS 181/group project/Data .csv")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#Summarize
summary(rawdat)
```
`
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
#Remove NAs
complete_dat <- na.omit(rawdat)

```


```{r}
#Check for duplicates
duplicated(complete_dat$Sample.ID)

```


```{r}
#Check if sample id is dup
complete_dat[!complete_dat$Patient.ID %in% complete_dat$Sample.ID, ]


```
```{r}
cols_nonunique <- c()

for(i in 1:ncol(complete_dat)){
  if(length(unique(complete_dat[,i])) <= 1){
    cols_nonunique[length(cols_nonunique) + 1] <- i
  }
}

print(cols_nonunique)

```

```{r}

remove_nonuniques <- complete_dat[,-cols_nonunique]

```


```{r}
#subset the data 
library(tidyverse)
age_tumor <- remove_nonuniques[c(4,11)]
age_tumor
summary(age_tumor$Diagnosis.Age)

age_tumor$agegrp <- ifelse(age_tumor$Diagnosis.Age >= 40 & age_tumor$Diagnosis.Age <=59, "40-59",
                  ifelse(age_tumor$Diagnosis.Age > 59 & age_tumor$Diagnosis.Age <= 69, "60-69",
                         ifelse(age_tumor$Diagnosis.Age > 69 & age_tumor$Diagnosis.Age <= 79, "70-79",
                                ifelse(age_tumor$Diagnosis.Age >79,"80+", NA))))

library(dplyr)
install.packages("gtsummary")
library(gtsummary)
age_tumor %>% 
  select(Tumor.Site, agegrp) %>%
  tbl_summary(by = agegrp, label = Tumor.Site ~ "Tumor Site") %>% add_p(test=everything()~"kruskal.test")
```

```{r}
freq <- age_tumor %>% group_by(agegrp,Tumor.Site) %>% summarise(Freq=n())
freq <- rename(freq, "Count of Tumors"=Freq)
freq_gt1 <-freq[which(freq$`Count of Tumors`>1),]
library(ggplot2)
library(viridis)
ggplot(freq_gt1, aes(agegrp,`Count of Tumors`)) +
  labs(x="Age Grouped by BJC paper levels", y="Frequency of Tumors") +
  geom_bar(aes(fill=`Tumor.Site`),stat="identity",position="dodge") +
  scale_fill_viridis_d(option  = "mako") 
  
```










